### 2014-01-08 reviews/response.md
Minor problems from reviewer 3

I really appreciate the thoughtful comments of all the reviewers. I've worked
hard to make the paper easier to understand and the reviewer's comments were
extremely useful.  A point-by-point summary of the comments and my rebuttal
is below. Minor comments fixed without needing further discussion are listed
at the end.

# Reviewer 1

> Details of writing: There is opportunity for improvement and for 
> strengthening the exposition at a local level – in sentences and phrases. 
> This reviewer has just read Wickham's 2011 paper in the same journal on
> "Split-Apply-Combine", and the writing in that paper is good and can serve 
> as a model for polishing the present paper. The differences between the two 
> are in the details, and a thorough editing and tightening of the present 
> manuscript would almost surely pay dividends. Without trying to be 
> comprehensive, I will illustrate the kinds of small changes (including 
> just fixing typos) which taken together would have a positive effect. (R1)

> p. 1 last full sentence: "The reorganization makes...because it conforms to 
> a standard that facilitates well an initial exploration and analysis of the 
> data; you don't..."

I'm not quite sure what the reviewer meant here, but I've rewritten the 
paragraph for clarity.

> p. 2 line 2: Is "reformulating" an approximation for "munging"?

Yes, but I think mungling better conveys the inelegance of this operation :)

> p. 3 bottom. Consider having a new bold subtitle as a guidepost for the 
> reader: Defining tidy data:

I've completely reorganised this section and it should be much more clear.

> p. 14 last paragraph. A difficulty is identified; does the author suggest 
> a solution? Reader isn't clear where we are left as we enter the Case Study.

> p. 21 down 13: might be clearer if "...an efficient equivalent to join." 
> – bolding or perhaps quotes.

I've removed those comments because I know no longer think that they a true.

> This reviewer notes (in a spirit of collegiality) that the best writers about 
> data analysis (from a previous generation) – Fred Mosteller, John Tukey,
> David Hoaglin, Paul Velleman, Frank Anscombe (as illustrations) – correctly 
> used data as plural form (and occasionally datum as singular). From that 
> perspective, "the data is ..." grates more than a little ... but 
> regrettably this is a battle that is now lost. Sigh... :) :)

# Reviewer 2

> While the topic is important and the message is clear, I found the tone of
> the paper to be too conversational. The author overuses contractions,
> ambiguous pronouns ("It's often said..."), and casually addresses the reader 
> in second-person form. (R2)

> The author also switches between the use of italics, boldface, quotes, and 
> the LaTeX verbatim environment constantly, which tends to be distracting 
> at times. (R2)
> - The author uses "colvars" in italics when first naming the terminology, and 
> then subsequently mentions the Pew dataset contains one colvar, without 
> italics. (R2)
> - The author does not use formatting on the column names in the caption for 
> Table 9 but does use \verbatim to reference column names in Table 10's 
> caption. (R2)

Thanks for pointing this out. Throughout the next I now use bold face 
consistently for the first definition of a new term, and verbatim for
variable names.

> The author overuses colons. In some cases, I found hyphens to be more 
> appropriate, and in others, I would think semicolons might be more suitable.

# Reviewer 3

> The author is obviously intimately familiar with the tools he's using in 
> this paper, and he should remember that the readers of this paper are not 
> nearly as familiar with them as he is. (R3)

> Section 2: "When using tables for communication, this so-called 
> "Alabama first" (Wainer 2000) ordering should be abandoned, and replaced 
> with an ordering based on a meaningful variable."
> 
> There doesn't seem to be any justification or explanation for this statement.
> (R3)

Removed.

> Section 3.1: Since most readers of this paper will want to apply these ideas 
> using R, I think it would be a good idea to show how an R datasets containing 
> the Pew and Billboard data would look, as well as the R code that would be 
> used to melt the datasets. This is especially true for the Billboard data, 
> because additional cleaning was performed. Similar comments apply to the 
> other datasets in the paper. (R3)

I think including the code in the paper would make it too long, but they
are all available at the paper website at https://github.com/hadley/tidy-data

> Section 3.5: I feel the examples in this section need to be explained in 
> more detail. The referenced external links contain a large number of files, 
> and it's not really clear which files are relevant to the discussion in this 
> section. (R3)

I've added a brief snippet of R code to illustrate how you might tackle this.

> Section 4.1: "Compare this to the difficulty of combining datasets stored in 
> arrays; these typically require painstaking alignment before matrix 
> operations can be used, and errors can be very hard to detect."
> Would anyone seriously suggest using matrices to store data which would be 
> combined with other data sets? The use of an index to combine multiple 
> datasets doesn't really seem to be related to tidy data. (R3)

Right, and I've now removed this sentence.

> Section 5: I was not able to download the data for the case study from 
> https://raw.github.com/hadley/tidy-data/master/case-study/deaths.rdata; 
> I recieved a 403 Forbidden error. (R3)

Smaller version that you can download now at 
https://raw.github.com/hadley/tidy-data/master/case-study/deaths.rds

> I think there should be a more thorough discussion explaining the goal of 
> analyzing this data set. I'm not familiar with the term "time course", and 
> the single sentence "The case study uses individual-level mortality data f
> rom Mexico, with the goal of finding causes of death that have notably 
> different time patterns within a day." really didn't clarify the goal 
> sufficiently for me. Perhaps an example of an unusual time course would help 
> to clarify things. (R3)

# Minor problems

> Abstract: "take in and take out" "input and output"

> p. 1 up 6: "subset" -> "component"

> p. 2 line 9: -> "an extension..."

> p. 2, 5 up from subtitle: "...techniques with real examples." -- could do 
> nicely.

> p. 2, 1 up from subtitle "...misses and what other approaches might be 
> fruitful to pursue."

> p. 3 2/3 down "...were were..."

> p. 10, up 2. Sentence is garbled.

> p. 11 bottom: (left) and (right) may not be the apt descriptors...

> p. 12, line 9: comma after file name?

> p. 12: This reviewer would prefer to avoid "hopefully" about 13 lines down. 
> More substantively, the sentence is garbled.

> p. 13 middle "...by the by preposition." Should second "by' be bolded? Or in 
> quotes" Same question arises again.

> p. 17 middle -> " diseases we work with have..."

> p. 19 last line "...seem like they should ..." -> "may"

> Section 4.3: I think readers of this paper might find examples from stata 
> more relevant than those from spss. (R3)

> The last sentence in the first paragraph of the discussion needs to be 
> rephrased for clarity.
> (R2)

> Some tables also have unreadable characters in them, such as the degree sign 
> in Table 12's top panel (row 6's artist). Same issue with Table 3's 5th 
> religion. Table 12's caption also describes there being a left and right 
> dataset, whereas the subtables are clearly placed above and below.
> (R2)

> - Abstract: "variables are stored in columns, observations in rows, and a 
> single type of..." -- this list isn't logically homogeneous.

> - p.5, 3.1 -- "The Pew Center is an American ... that collects data on 
> attitudes to topics RANGING from religion to the internet"

> - "Billboard" should always be capitalized as it is a proper noun.

> - p.13, section 4.2 - "Tidy visualization tools ONLY NEED to be..."

> - p.16, 4-th line from bottom: "Next, we work out THE overall..."

> - p.16, 2nd line from the bottom: "Then finally, WE join.."

> - p.19, "The causes of death fall INTO three main groups: ... " There should 
> be a hyphen between "transportation" and "related".

> - p. 19, 2nd paragraph of discussion: "This makes it easy"; remove "is".

> - p. 21, last sentence of first paragraph: "and A BETTER KNOWLEDGE OF how 
> we can best design tools..."

> - p. 21, last paragraph before Acknowledgements "verifying experimental 
> design, AND filling in..."

> - Author information, last page "Adjunct ASSISTANT Professor"